---
title: "lending_club_projecty_part1"
author: "Haichuan Du"
date: "5/21/2020"
output: pdf_document
---
# Read data

```{r setup, include=FALSE}
loan <- read.csv("/Users/haidu/Desktop/ds501/bittiger/lendingClu_project/loan-data_project_1/loan.csv", stringsAsFactors = FALSE)
loanT <- loan
head(loan)
```

Check dimension
```{r}
dim(loan)
```

check data format
```{r}
str(loan)
```

Calculate the number of na values for each column.    
```{r}
#seperate columns with over 80% NA values
num.NA <- sort(colSums(is.na(loan)), decreasing = TRUE)
remain.col <- names(num.NA)[num.NA < 0.8 * dim(loan)[1]]
delete.col <- names(num.NA)[num.NA >= 0.8 * dim(loan)[1]]
```

# exploratory data analysis (EDA) (What factor will influence the interest rate?) 

```{r}
{
  plot(density(loan$int_rate), main = "Density plot of interest rate", xlab = "Interest rate")
  abline(v = mean(loan$int_rate), lty = 2, col = 2)
  abline(v = median(loan$int_rate), lty=2, col=3)
  abline(v = quantile(loan$int_rate, 0.2), lty=2, col=4)
  abline(v = quantile(loan$int_rate, 0.8), lty=2, col=6)
  legend("topright", c("mean", "median", "20% quantile", "80% quantile"),
         col = c(2, 3, 4, 6), lty = 2, title = "Interest rate statistics")
}

```
The distribution is a bit right skew. In the future, if we want to build model with interest rate then it is better to use square root to adjust the skewness. Next I want to explore the correlation between interest rate with other numeric variables. However, we know some features are with high number of NA values. So I want to remove those features with over 80% NA values temperarily.  

```{r}
library(corrplot)
loan <- loan[, remain.col]
#select numerical features' name
num.feature <- names(loan[, sapply(loan, is.numeric)])
#select char features' name
char.feature <- names(loan[, sapply(loan, is.character)])
#calculate the correlationship between int_rate and other numerical features
correlation <- cor(loan$int_rate, loan[, num.feature], use = "pairwise.complete.obs")
#sort the value
correlation <- correlation[, order(correlation[1,], decreasing = TRUE)] 
#remove correlation with itself and correlation with police_code
correlation <- correlation[2:33] #刚好排除第一个和最后一个
```
https://www.guru99.com/r-apply-sapply-tapply.html 

```{r}
{
  barplot(correlation, main = 'Correlation between int_rate with other numerical features', ylab = "Correlation", las = 3, cex.names = 0.7, col = ifelse(correlation > 0.15, "red", "blue"))
  legend("topright", c('top five correlated feature', 'other numerical feature'), fill = c("red", "blue") )
}
```
Based on correlation calculation, the top five predictive numerical features for int_rate are "total_rec_int", "revol_util", "inq_last_6mths", "total_pymnt_inv" and "total_pymnt".

Next step is to explore the top five influencial category features.  

```{r}
library(ggplot2)
ggplot(data = loan, aes(term, int_rate, fill = term)) +
  geom_boxplot(outlier.colour = "blue") +
  labs(title = "Box plot of Interest rate against grade") +
  theme(plot.title = element_text(hjus = 0.5)) #居中标题
```

Student's t-test assumes that the two populations have normal distributions with equal variances.
Welch's t-test is designed for unequal variances, but the assumption of normality is maintained,
我的method1的公示表第4个, l2个sample不一定要follow normal 但是sample size 需要大于30一般来说
```{r}
t.test(subset(loan, term == " 36 months")$int_rate, 
       subset(loan, term == " 60 months")$int_rate,
       conf.level = 0.95, mu = 0, alternative = "two.side", paired = FALSE, var.equal = FALSE)
```
注意subset的用法， subset(loan, term == " 36 months") 输出来的是datafram 和loan 一样只不过是只包含了36 month,
There is a significant difference between different term. Therefore, term can be used a predictor for interest rate.   

```{r}
ggplot(loan, aes(grade, int_rate, fill = grade)) +
  geom_boxplot(outlier.colour = "blue") +
  labs(title = "Box plot of Interest rate against grade") +
  theme(plot.title = element_text(hjust = 0.5))
```
There is a clear <<<<linear>>>> trend between interest rate and grade. Therefore, grade can be used a predictor for interest rate. However, based on the description of grade. It is assigned by the Lending Club. Therefore, I probably don't have this feature in advance. If a client is a return user and I can definitely use this information.     

```{r}
ggplot(data=loan, aes(verification_status,int_rate,fill=verification_status)) +
  geom_boxplot(outlier.color = "blue") +
  labs(title="Box plot of interest rate against verification_status") + 
  theme(plot.title = element_text(hjust = 0.5))

```

```{r}
t.test(subset(loan, verification_status == 'Verified')$int_rate, subset(loan, verification_status == 'Source Verified')$int_rate, conf.level = 0.95, mu = 0, alternative = "two.sided", paired = FALSE, var.equal = FALSE)

t.test(subset(loan, verification_status == 'Not Verified')$int_rate, subset(loan, verification_status == 'Source Verified')$int_rate, conf.level = 0.95, mu = 0, alternative = "two.sided", paired = FALSE, var.equal = FALSE)

t.test(subset(loan, verification_status == 'Not Verified')$int_rate, subset(loan, verification_status == 'Verified')$int_rate, conf.level = 0.95, mu = 0, alternative = "two.sided", paired = FALSE, var.equal = FALSE)
```

There is a significant difference between different verification_status Therefore, verification_status can be used a predictor for interest rate.       

```{r}
ggplot(data=loan, aes(pymnt_plan,int_rate,fill=pymnt_plan))+
  geom_boxplot(outlier.color = "blue")+
  labs(title="Box plot of interest rate against pymnt_plan") + 
  theme(plot.title = element_text(hjust = 0.5))
```

```{r}
t.test(subset(loan, pymnt_plan == 'n')$int_rate, subset(loan, pymnt_plan == 'y')$int_rate, conf.level = 0.95, mu = 0, alternative = "two.sided", paired = FALSE, var.equal = FALSE)

```

There is a significant difference between different pymnt_plan. Therefore, pymnt_plan. can be used a predictor for interest rate.        

```{r}
ggplot(data=loan, aes(application_type,int_rate,fill=application_type))+
  geom_boxplot(outlier.color = "blue")+
  labs(title="Box plot of interest rate against application_type") +
  theme( plot.title = element_text(hjust = 0.5))

```

```{r}
t.test(subset(loan, application_type == 'INDIVIDUAL')$int_rate, subset(loan, application_type == 'JOINT')$int_rate, conf.level = 0.95, mu = 0, alternative = "two.sided", paired = FALSE, var.equal = FALSE)

```

There is a significant difference between different application_type. Therefore, application_type can be used a predictor for interest rate. In conclustion, these five category variables are influenciable: "term", "grade", "verification_status", "pymnt_plan" and "application_type". Besides I am also curious about how interest rate vary with space and time.       

```{r}
library(dplyr)
suppressPackageStartupMessages(library(maps))
loan$region <- loan$addr_state
typeof(loan$region)
loan$region <- as.factor(loan$region)
levels(loan$region)<- c("alaska", "alabama","arkansas", "arizona", "california","colorado","connecticut","district of columbia","delaware","florida","georgia","hawaii","iowa","idaho","illinois","indiana","kansas","kentucky","louisiana","massachusetts","maryland","maine","michigan","minnesota","missouri","mississippi","montana","north carolina","north dakota","nebraska","new hampshire","new jersey","new mexico","nevada","new york","ohio","oklahoma","oregon","pennsylvania","rhode island","south carolina","south dakota","tennessee","texas","utah","virginia","vermont","washington","wisconsin","west virginia","wyoming")

all_states <- map_data("state")
state_by_rate <- loan %>% group_by(region) %>% 
                  summarise(value = mean(int_rate, na.rm = TRUE))
state_by_rate$region <- as.character(state_by_rate$region)

Total <- merge(all_states, state_by_rate, by="region")

p <- ggplot()
p <- p + geom_polygon(data=Total, aes(x=long, y=lat, group = group, fill=Total$value),colour="white"
      ) + scale_fill_continuous(low = "skyblue", high = "darkblue", guide="colorbar")
P1 <- p + theme_bw()  + labs(fill = "Interest rate" 
                            ,title = "Heat map of interest rate in all states", x="", y="")
P1 + scale_y_continuous(breaks=c()) + scale_x_continuous(breaks=c()) + theme(panel.border =  element_blank(), plot.title = element_text(hjust = 0.5))

```

# EDA_part_2 (What are the distribution of loan status?)         

```{r}
library(DescTools)
Desc(loan$loan_status, plotit = TRUE, main = "Loan Status")
```

See how does loan_status affact interest rate.    

```{r}
ggplot(data = loan, aes(loan_status, int_rate), las=2) +
  geom_boxplot(aes(fill = loan_status)) +
  labs(title = "Interest rate by loan_status", x = "Loan_status",y = "Interest rate") +
  theme(axis.text.x = element_blank(), plot.title = element_text(hjust = 0.5))
```

The majority status is "Current". The finished loan can be grouped into "Fully Paid" and "Charged off" or "Late payment".       

# EDA_part_3 (What are the purpose of applying a loan with Lending Club ?)       


```{r}
Desc(loan$purpose, main = "Loan purposes", plotit = TRUE)

```

I am also curious about how interest rate vary with purpose.      

```{r}
ggplot(data=loan,aes(purpose,int_rate), las=2) +
  geom_boxplot(aes(fill=purpose))+
  labs(list(title = "Interest rate by purpose",x = "purpose of Loan",y = "Interest rate"))+
  theme(axis.text.x = element_blank(), axis.title.x = element_blank(), axis.ticks.x = element_blank(), plot.title = element_text(hjust = 0.5))
```




